---
title: "Black Communities, Crime, and Police"
author: "Aser Atawya"
date: "5/8/2022"
output: html_document
---

## Introduction
  The BLM movement and the demand for equal justice for Black people have been growing exponentially since the tragic incident of George Floyd. Inspired by the BLM, I believe the scientific study and analysis of the data sets recording police interactions with the public and the crime rates in Black communities can support Black activism by highlighting the racial discrimination against Black people. This project aims to answer three important questions: What is the racial distribution of the victims and the perpetrators of homicide crimes in the U.S? What is the influence of race on the victims of police shootings?  Is there a racial discrimination against Black people when it comes to the drivers stopped by police and the duration of their stop? To answer the aforementioned questions, I plan to analyze four data sets. The first data set has a record for all police shootings since 2000, including the race and gender of the victim. The second data set has a detailed record for every homicide case in the US from 1980 to 2014, including gender and race of both the victim and perpetrator.  The third data set has a record of all police stops in Rhode Island from 2005 to 2015. The police stops are limited to Rhode Island only due to the size of the data; there are 200 million police stops recorded across the U.S, so analyzing such data sets is beyond the scope of this project. The fourth data set is a census of every county and state in the U.S by race and ethnicity. 
  
  The project found that there is a disproportionate ratio between the percentage of Black perpetrators of homicide crimes in the U.S and the percentage of Black people in the U.S population. Similarly, there is an even more skewed ratio between the percentage of Black victims of homicide crimes in the U.S and the proportion of Black people in the U.S population. Thus, the answer to the first question shows that Black communities are considerably vulnerable to crimes. The findings, also, showed that over one fifth of police shootings victims were Black even though the proportion of Black people in the U.S population is less than 15%.  Furthermore, not only were the proportion of Black drivers stopped by the police in Rhode Island far more than the percentage of Black people in the state but also the Black drivers were stopped for longer duration compared to White drivers. The answers to the second and third questions underlines the unjust treatment of Black people by the police across the U.S.
    
    
    
## Data Source

US Police Shootings: https://fatalencounters.org/
The data was first crowdsourced and validated by comparing the data to public media reports. Then, the organization hired paid researchers to collect data from reliable sources, such as the LA Times Homicide report. Furthermore, the organization collected data directly from law enforcement agencies, requesting over 2300 public records of state, federal and local law enforcement agencies. 

Homicide Reports, 1980-2014: https://www.kaggle.com/datasets/murderaccountability/homicide-reports
The data was collected directly from the FBI master files of the Uniform Crime Report’s "Return A", which has a summary of the homicide crimes in the U.S that ended with the arrest of the perpetrator, and the FBI's original Supplementary Homicide Report, which records the details of every homicide crime in the U.S since 1976.

Stanford Open Policing Project: https://www.kaggle.com/datasets/faressayah/stanford-open-policing-project
The data was collected by Stanford Computational Policy Lab in collaboration with Big Local News by making public record requests of state and local police departments.

Population by Race and Ethnicity: https://www.statsamerica.org/downloads/default.aspx
The source of the data is the U.S Census Bureau, which collects data from federal, state, and local governments; in addition, some of the data is collected directly from citizens through surveys. Data sources include federal agencies, such as the IRS, and commercial businesses working on data collection. The data on race and ethnicity are based on self-identification.

## Data Ethics

The analysis of the data brings light to the racial discrimination against Black communities in multiple areas, police interactions and homicide rates. The provision of accurate data regarding this matter is the basis of the intervention to achieve social efficiency, which is an ethically important gain. The analysis of the racial distribution of police shootings and police stops should impel police departments across the U.S to adopt better institutional policies to combat this racial bias, which is an ethically important gain. Highlighting the skewed ratio between the percentage of Black people in the population and the percentage of Black homicide victims and perpetrators develops our human understanding of the issue and lets us act more wisely to provide more safety to Black communities, which is another ethically important gain.

 The data collected by the U.S Census Bureau, Stanford Open Policing Project, and the Murder Accountability Project avoids any harm to privacy and security because the data is anonymous in contrast to the data of police shootings, which has the full name of the victims. The data of the police shootings can be used to link the victims of police shootings to their family members using the full name of the victim; if the data was used in such a manner, the family members may be treated unfairly because they are linked to potentially dangerous criminals who were shot by the police, which is an ethically significant harm. Furthermore, this can lead the relatives of the victims to experience reputational or emotional harm, which are ethically significant harms.




## Data Import, Cleaning and Tidying, and Exploration

```{r, message = FALSE}
#calls all the needed packages
library(tidyverse)
library(ggplot2)
library(modelr)
library("scatterplot3d")
options(na.action = na.warn)
```

### - Data Import
```{r}
#imports the four databases
police_shootings <- read_csv("Police_Shootings.csv")
police_stops <- read_csv("Police_Stops.csv")
homicide <- read_csv("Homicide.csv")
US_race <- read_csv("Population by Race.csv")
police_shootings
police_stops
homicide
US_race
```

### - Data Tidying, Cleaning, and Transformation


```{r}
#calculates the percentage of each race in the U.S population 
USA_Race <- US_race %>% filter(Description == "U.S.") %>% mutate(
  Black_Proportion  = `Black Alone`*100/ `Total Population`, White_Proportion = `White Alone`*100/ `Total Population`, 
  Asian_Proportion = `Asian Alone`*100/ `Total Population`, Hispanic_Proportion = Hispanic*100/ `Total Population`,
  Native_Proportion = `American Indian or Alaskan Native`  *100/ `Total Population`, 
  Hawaiian_Proportion = `Hawaiian or Pacific Islander Alone` *100/ `Total Population`
)
USA_Race 
```

```{r}
#checks the values in the Race and Race with imputations columns
police_shootings %>% distinct(Race)
police_shootings %>% distinct(`Race with imputations`)
```

  
```{r}
#removes the different spellings of the same race
police_shootings_clean <- mutate(police_shootings, 
Race = ifelse(Race %in% c("European-American/White", "european-American/White", "European-American/European-American/White"), "White", Race),
Race = ifelse(Race %in% c("African-American/Black", "African-American/Black African-American/Black Not imputed"), "Black", Race),

#if imputation probability is more than 50% and the race is unspecified, assign the imputed race
Race = ifelse( Race == "Race unspecified" & `Race with imputations` %in% c("African-American/Black", "African-American/Black African-American/Black Not imputed") & `Imputation probability` > 0.5, "Black", Race), 
Race = ifelse( Race == "Race unspecified" & `Race with imputations` %in% c("European-American/White", "european-American/White", "European-American/European-American/White") & `Imputation probability` > 0.5, "White", Race),
) %>% 
  
#removes one row with invalid Race Value
filter( Race != "Christopher Anthony Alexander", !is.na(Race))

police_shootings_clean
```



```{r}
#parses date and chooses only the relevant columns
police_shootings_clean_2 <- police_shootings_clean %>% mutate(Date = parse_date(`Date of injury resulting in death (month/day/year)`,"%m/%d/%Y"), Year = str_sub(Date, 1, 4)) %>% select(Year, Gender, Race, `Armed/Unarmed`)

#calculates the total police shootings each year
police_shootings_by_year <- police_shootings_clean_2 %>% group_by(Year)  %>% summarize(n())
#calculates the police shootings by race each year
police_shootings_by_race <- police_shootings_clean_2 %>% group_by(Year)  %>%  count(Race) %>% pivot_wider(names_from = Race, values_from = n) 

#joins the two tibbles and calculates the racial distribution of the people shot by police 
police_shootings_by_race_and_year <-  full_join(police_shootings_by_year, police_shootings_by_race, by = "Year") %>% rename( Total = `n()`   ) %>% mutate(Year = parse_double(Year), Asian_Percentage = `Asian/Pacific Islander`*100/ Total, Black_Percentage = Black*100/ Total, Hispanic_Percentage = `Hispanic/Latino` *100/ Total, Native_Percentage =  `Native American/Alaskan`*100/ Total, Unspecified_Percentage =`Race unspecified`*100/ Total, White_Percentage = White*100/ Total, Middle_Eastern_Percentage =  `Middle Eastern`*100/ Total) 

#joins the tibble with the racial distribution of the people shot by police and the tibble with the racial distribution of the U.S population
police_shootings_And_US_Race <- inner_join(police_shootings_by_race_and_year, USA_Race, by = "Year") 
police_shootings_And_US_Race
```

```{r}
#calculates the total homicide crimes each year
Homicide_by_year <- homicide %>% group_by(Year)  %>% summarize(n())

#calculates the total homicide perpetrators each year by race
Homicide_by_perpetrator_race <- homicide%>% group_by(Year)  %>%  count(`Perpetrator Race`) %>% pivot_wider(names_from = `Perpetrator Race`, values_from = n) 

#calculates the total homicide victims each year by race and rename the columns to avoid overlapping with the tibble Homicide_by_perpetrator_race
Homicide_by_victim_race <- homicide%>% group_by(Year)  %>%  count(`Victim Race`) %>% pivot_wider(names_from = `Victim Race`, values_from = n) %>% rename(  `Asian/Pacific Islander_victim` = `Asian/Pacific Islander`, Black_victim = Black  , `Native American/Alaska Native_victim` = `Native American/Alaska Native` ,  Unknown_victim = Unknown, White_victim = White)

#joins all the three tibbles and calculates the racial distribution of homicide victims and perpetrator
Homicide_by_race_and_year <-  full_join(Homicide_by_year, Homicide_by_perpetrator_race , by = "Year") %>%  full_join(Homicide_by_victim_race , by = "Year") %>% rename( Total = `n()`   ) %>% mutate( Asian_Percentage = `Asian/Pacific Islander`*100/ Total, Black_Percentage = Black*100/ Total,  Native_Percentage =  `Native American/Alaska Native`*100/ Total, Unspecified_Percentage =`Unknown`*100/ Total, White_Percentage = White*100/ Total, Asian_Percentage_victim = `Asian/Pacific Islander_victim`*100/ Total, Black_Percentage_victim = Black_victim*100/ Total,  Native_Percentage_victim =  `Native American/Alaska Native_victim`*100/ Total, Unspecified_Percentage_victim =`Unknown_victim`*100/ Total, White_Percentage_victim = White_victim *100/ Total) 

#joins the tibble with the racial distribution of homicide and the tibble with the racial distribution of the U.S population
Homicide_And_US_Race <- inner_join(Homicide_by_race_and_year, USA_Race, by = "Year") 
Homicide_And_US_Race
```


```{r}
#calculates the percentage of each race in the population of Rhode Island
Rhode_Island_Race <- US_race %>% filter (Description == "Rhode Island") %>% mutate(
  Black_Proportion = `Black Alone`*100/ `Total Population`, White_Proportion = `White Alone`*100/ `Total Population`, 
  Asian_Proportion = `Asian Alone`*100/ `Total Population`, Hispanic_Proportion = Hispanic*100/ `Total Population`,
  Native_Proportion = `American Indian or Alaskan Native`  *100/ `Total Population`, 
  Hawaiian_Proportion = `Hawaiian or Pacific Islander Alone` *100/ `Total Population`
) %>%
  # rename Hispanic to avoid the automatic renaming when joining this tibble in the next code chunk
  rename(Hispanic_race = Hispanic)
Rhode_Island_Race 
```



```{r}
#creates a Year column
police_stops_2 <- police_stops %>% mutate(Year = str_sub(stop_date, 1, 4))

#calculates the total police stops each year in Rhode Island
police_stops_2_by_year <-police_stops_2 %>% group_by(Year)  %>% summarize(n())

#calculates the police stops by race each year in Rhode Island
police_stops_2_by_race <- police_stops_2 %>% group_by(Year)  %>%  count(driver_race) %>% pivot_wider(names_from = driver_race, values_from = n) 

#joins the two tibbles and calculates the racial distribution of the people stopped by police in Rhode Island 
police_stops_by_race_and_year <-  full_join(police_stops_2_by_race, police_stops_2_by_year, by = "Year") %>% rename( Total = `n()`   ) %>% mutate(Year = parse_double(Year), Asian_Percentage = Asian*100/ Total, Black_Percentage = Black*100/ Total, Hispanic_Percentage = Hispanic *100/ Total,  White_Percentage = White*100/ Total) 

#joins the tibble with the racial distribution of the people stopped by the police in Rhode Island and the tibble with the racial distribution of Rhode Island population
police_stops_And_Rhode_Island_Race <- inner_join(police_stops_by_race_and_year, Rhode_Island_Race, by = "Year") 
police_stops_And_Rhode_Island_Race
```
```{r}
#checks all the values of the stop_duration column
police_stops_2 %>% distinct(stop_duration)
```

```{r}
#groups the data by stop_duration and counts the people stopped by their race and avoids the irrelevant values of stop_duration
police_stops_duration_by_race <- police_stops_2 %>% filter(stop_duration != 2 & stop_duration != 1) %>% group_by(stop_duration)  %>%  count(driver_race) %>% pivot_wider(names_from = driver_race, values_from = n) 
police_stops_duration_by_race

#groups the data by race and counts the people stopped by their stop_duration and avoids the irrelevant values of stop_duration
police_stops_duration_by_race2 <- police_stops_2 %>% filter(stop_duration != 2 & stop_duration != 1) %>% group_by(driver_race)  %>%  count(stop_duration) %>% pivot_wider(names_from = stop_duration, values_from = n) 
police_stops_duration_by_race2
```



```{r}
#removes the irrelevant values in the stop_duration column
police_stops_filtered <- police_stops_2 %>% filter(stop_duration =="0-15 Min" | stop_duration =="16-30 Min" | stop_duration =="30+ Min")
police_stops_filtered
```
  


```{r}
#checks the values of the Description column of US_race
US_race %>% distinct(Description)
```
 

```{r}
#finds the U.S states where Black people are a majority
#the data was filtered to keep only the 50 states and D.C
US_race_filtered <- US_race %>% filter(!str_detect(Description, ","), Description != "U.S.", Description != "Puerto Rico") 
US_race_filtered %>% mutate(majority = pmax(US_race_filtered $`White Alone`, US_race_filtered $`Black Alone`, US_race_filtered $`American Indian or Alaskan Native`, US_race_filtered $`Asian Alone`, US_race_filtered $`Hawaiian or Pacific Islander Alone`, US_race_filtered $`Two or More Races`, na.rm = TRUE), black_majority = ifelse(majority == US_race_filtered$`Black Alone`, 1, 0)) %>% select(majority, black_majority, everything()) %>% filter(black_majority == 1) %>% distinct(Description)
```
None of the 50 U.S states has a Black majority. Only the District of Columbia has a black majority. 



### - Data Visualization 

```{r}
#column plot of the percentage of police shooting victims who are Black with a red line showing the percentage of Black people in the population

ggplot(data = police_shootings_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Black_Percentage), fill = "Black") + labs(title="Percentage of Police Shooting Victims \n Who Are Black", x = "Year", y = "Black Percentage of Police Shootings") + geom_line(aes(x = Year, y = Black_Proportion), color = "Red")
```

```{r}
#column plot of the percentage of police shooting victims who are White with a red line showing the percentage of White people in the population

ggplot(data = police_shootings_And_US_Race ) + geom_col(mapping = aes(x = Year, y = White_Percentage), fill = "Black") + labs(title="Percentage of Police Shooting Victims \n Who Are White", x = "Year", y = "White Percentage of Police Shootings") + geom_line(aes(x = Year, y = White_Proportion), color = "Red")
```

```{r}
#column plot of the percentage of police shooting victims who are Hispanic with a red line showing the percentage of Hispanic people in the population

ggplot(data = police_shootings_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Hispanic_Percentage), fill = "Black") + labs(title="Percentage of Police Shooting Victims \n Who Are Hispanic", x = "Year", y = "Hispanic Percentage of Police Shootings") + geom_line(aes(x = Year, y = Hispanic_Proportion), color = "Red")
```


```{r}
#column plot of the percentage of police shooting victims who are Asian with a red line showing the percentage of Asian people in the population

ggplot(data = police_shootings_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Asian_Percentage), fill = "Black") + labs(title="Percentage of Police Shooting Victims \n Who Are Asian", x = "Year", y = "Asian Percentage of Police Shootings") + geom_line(aes(x = Year, y = Asian_Proportion), color = "Red")
```

```{r}
#column plot of the percentage of police shooting victims who are Native Americans with a red line showing the percentage of Native Americans people in the population

ggplot(data = police_shootings_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Native_Percentage), fill = "Black") + labs(title="Percentage of Police Shooting Victims \n Who Are Native Americans", x = "Year", y = "Native Americans' Percentage of Police Shootings") + geom_line(aes(x = Year, y = Native_Proportion), color = "Red")
```






```{r}
#column plot of the percentage of homicide perpetrators who are Black with a red line showing the percentage of Black people in the population

ggplot(data = Homicide_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Black_Percentage), fill = "Black") + labs(title="Percentage of Homicide Perpetrators \n Who Are Black", x = "Year", y = "Black Percentage of Homicide Perpetrators") + geom_line(aes(x = Year, y = Black_Proportion), color = "Red")
```

```{r}
#column plot of the percentage of homicide victims who are Black with a red line showing the percentage of Black people in the population
ggplot(data = Homicide_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Black_Percentage_victim), fill = "Black") + labs(title="Percentage of Homicide Victims \n Who Are Black", x = "Year", y = "Black Percentage of Homicide Victims") + geom_line(aes(x = Year, y = Black_Proportion), color = "Red")
```


```{r}
#3dplot of the percentage of homicide perpetrators who are Black and the percentage of homicide victims who are Black with a plane showing the average percentage of Black people in the population

s <- scatterplot3d(x = Homicide_And_US_Race$Year, y = Homicide_And_US_Race$Black_Percentage_victim, z = Homicide_And_US_Race$Black_Percentage, ylim = c(0,100), zlim = c(0,100),type = "h", angle = 15, highlight.3d = TRUE, xlab = "Year", ylab = "Black Percentage of Homicide Victims", zlab = "Black Percentage of Homicide Perpetrators", main="Black Homicide Victims and Perpetrators")
s$plane3d(mean(Homicide_And_US_Race$Black_Proportion),0,0,lty="solid")

```




```{r}
#column plot of the percentage of homicide perpetrators who are White with a red line showing the percentage of White people in the population

ggplot(data = Homicide_And_US_Race ) + geom_col(mapping = aes(x = Year, y = White_Percentage), fill = "Black") + labs(title="Percentage of Homicide Perpetrators \n Who Are White", x = "Year", y = "White Percentage of Homicide Perpetrators") + geom_line(aes(x = Year, y = White_Proportion), color = "Red")
```

```{r}
#column plot of the percentage of homicide victims who are White with a red line showing the percentage of White people in the population

ggplot(data = Homicide_And_US_Race ) + geom_col(mapping = aes(x = Year, y = White_Percentage_victim), fill = "Black") + labs(title="Percentage of Homicide Victims \n Who Are White", x = "Year", y = "White Percentage of Homicide Victims") + geom_line(aes(x = Year, y = White_Proportion), color = "Red")
```


```{r}
#3dplot of the percentage of homicide perpetrators who are White and the percentage of homicide victims who are White with a plane showing the average percentage of White people in the population

s <- scatterplot3d(x = Homicide_And_US_Race$Year, y = Homicide_And_US_Race$White_Percentage_victim, z = Homicide_And_US_Race$White_Percentage, ylim = c(0,100), zlim = c(0,100),type = "h", angle = 15, highlight.3d = TRUE, xlab = "Year", ylab = "White Percentage of Homicide Victims", zlab = "White Percentage of Homicide Perpetrators", main="White Homicide Victims and Perpetrators")
s$plane3d(mean(Homicide_And_US_Race$White_Proportion),0,0,lty="solid")

```

```{r}
#column plot of the percentage of homicide perpetrators who are Asian with a red line showing the percentage of Asians in the population

ggplot(data = Homicide_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Asian_Percentage), fill = "Black") + labs(title="Percentage of Homicide Perpetrators \n Who Are Asian/Pacific Islander", x = "Year", y = "Asian/Pacific Islander Percentage of Homicide Perpetrators") + geom_line(aes(x = Year, y = Asian_Proportion), color = "Red")
```

```{r}
#column plot of the percentage of homicide victims who are Asian with a red line showing the percentage of Asians in the population

ggplot(data = Homicide_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Asian_Percentage_victim), fill = "Black") + labs(title="Percentage of Homicide Victims \n Who Are Asian/Pacific Islander", x = "Year", y = "Asian/Pacific Islander Percentage of Homicide Victims") + geom_line(aes(x = Year, y = Asian_Proportion), color = "Red")
```


```{r}
#3dplot of the percentage of homicide perpetrators who are Asian and the percentage of homicide victims who are Asian with a plane showing the average percentage of Asians in the population

s <- scatterplot3d(x = Homicide_And_US_Race$Year, y = Homicide_And_US_Race$Asian_Percentage_victim, z = Homicide_And_US_Race$Asian_Percentage, ylim = c(0,10), zlim = c(0,10),type = "h", angle = 15, highlight.3d = TRUE, xlab = "Year", ylab = "Asian Percentage of Homicide Victims", zlab = "Asian Percentage of Homicide Perpetrators", main="Asian/Pacific Islander Homicide Victims and Perpetrators")
s$plane3d(mean(Homicide_And_US_Race$Asian_Proportion),0,0,lty="solid")

```



```{r}
#column plot of the percentage of homicide perpetrators who are Native Americans with a red line showing the percentage of Native Americans in the population

ggplot(data = Homicide_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Native_Percentage), fill = "Black") + labs(title="Percentage of Homicide Perpetrators \n Who Are Native American/Alaska Native", x = "Year", y = "Native American/Alaska Native Percentage of Homicide Perpetrators") + geom_line(aes(x = Year, y = Native_Proportion), color = "Red")
```

```{r}
#column plot of the percentage of homicide victims who are Native Americans with a red line showing the percentage of Native Americans in the population

ggplot(data = Homicide_And_US_Race ) + geom_col(mapping = aes(x = Year, y = Native_Percentage_victim), fill = "Black") + labs(title="Percentage of Homicide Victims \n Who Are Native American/Alaska Native", x = "Year", y = "Native American/Alaska Native Percentage of Homicide Victims") + geom_line(aes(x = Year, y = Native_Proportion), color = "Red")
```


```{r}
#3dplot of the percentage of homicide perpetrators who are Native Americans and the percentage of homicide victims who are Native Americans with a plane showing the average percentage of Native Americans in the population

s <- scatterplot3d(x = Homicide_And_US_Race$Year, y = Homicide_And_US_Race$Native_Percentage_victim, z = Homicide_And_US_Race$Native_Percentage, ylim = c(0,3), zlim = c(0,3),type = "h", angle = 15, highlight.3d = TRUE, xlab = "Year", ylab = "Native American % of Homicide Victims", zlab = "Native American % of Homicide Perpetrators", main="Native Americans Homicide Victims and Perpetrators")
s$plane3d(mean(Homicide_And_US_Race$Native_Proportion),0,0,lty="solid")

```







```{r}
#column plot of the percentage of police stops who are Black with a red line showing the percentage of Black people in the population of Rhode Island

ggplot(data = police_stops_And_Rhode_Island_Race) + geom_col(mapping = aes(x = Year, y = Black_Percentage), fill = "Black") + labs(title="Percentage of Police Stops \n Wherein the driver is Black", x = "Year", y = "Black Percentage of Police Stops") + geom_line(aes(x = Year, y = Black_Proportion), color = "Red")
```


```{r}
#column plot of the percentage of police stops who are White with a red line showing the percentage of White people in the population of Rhode Island

ggplot(data = police_stops_And_Rhode_Island_Race) + geom_col(mapping = aes(x = Year, y = White_Percentage), fill = "Black") + labs(title="Percentage of Police Stops \n Wherein the driver is White", x = "Year", y = "White Percentage of Police Stops") + geom_line(aes(x = Year, y = White_Proportion), color = "Red")
```
```{r}
#column plot of the percentage of police stops who are Hispanic with a red line showing the percentage of Hispanics in the population of Rhode Island

ggplot(data = police_stops_And_Rhode_Island_Race) + geom_col(mapping = aes(x = Year, y = Hispanic_Percentage), fill = "Black") + labs(title="Percentage of Police Stops \n Wherein the driver is Hispanic", x = "Year", y = "Hispanic Percentage of Police Stops") + geom_line(aes(x = Year, y = Hispanic_Proportion), color = "Red")
```


```{r}
#column plot of the percentage of police stops who are Asian with a red line showing the percentage of Asian in the population of Rhode Island

ggplot(data = police_stops_And_Rhode_Island_Race) + geom_col(mapping = aes(x = Year, y = Asian_Percentage), fill = "Black") + labs(title="Percentage of Police Stops \n Wherein the driver is Asian", x = "Year", y = "Asian Percentage of Police Stops") + geom_line(aes(x = Year, y = Asian_Proportion), color = "Red")
```


```{r}
#bar plot shows the racial distribution for each stop_duration 
ggplot(police_stops_filtered ) + geom_bar(mapping = aes(x= stop_duration, fill = driver_race), position = "fill")  + labs(title="The Racial Distribution \n for Each Stop Duration", x = "Duration of Stop", y = "Race of the Driver")
#bar plot shows the distribution of stop_duration for each race
ggplot(police_stops_filtered ) + geom_bar(mapping = aes(x= driver_race, fill = stop_duration), position = "fill")  + labs(title="The Distribution of Stop Duration\n for Each Race", x = "Race of the Driver", y = "Duration of Stop")
```
  
  
  
  
  
  
### - Data Modelling 


```{r}
#creates training data out of the original data after selecting the even years
Homicide_And_US_Race_training <- Homicide_And_US_Race %>% filter(Year %% 2 == 0)

#finds whether Black or White is the highest possibility for the race of a homicide victim
mean(Homicide_And_US_Race_training$Black_Percentage_victim)
mean(Homicide_And_US_Race_training$White_Percentage_victim)
```
The average percentage of a homicide victim being black is higher than any other race. Thus, the baseline model is that for every state, the majority of homicide victims are black.




```{r}
#creates test data by selecting the odd years and grouping the homicide victims by states
homicide_test <- homicide %>% filter(Year %% 2 == 1) %>% group_by(State)  %>%  count(`Victim Race`) %>% pivot_wider(names_from = `Victim Race`, values_from = n) %>% rename(  `Asian/Pacific Islander_victim` = `Asian/Pacific Islander`, Black_victim = Black  , `Native American/Alaska Native_victim` = `Native American/Alaska Native` ,  Unknown_victim = Unknown, White_victim = White) %>% mutate( Total = sum(`Asian/Pacific Islander_victim`, Black_victim, `Native American/Alaska Native_victim` ,Unknown_victim, White_victim, na.rm = TRUE), Asian_Percentage_victim = `Asian/Pacific Islander_victim`*100/ Total, Black_Percentage_victim = Black_victim*100/ Total,  Native_Percentage_victim =  `Native American/Alaska Native_victim`*100/ Total, Unspecified_Percentage_victim =`Unknown_victim`*100/ Total, White_Percentage_victim = White_victim *100/ Total)

#adds the result of the test to the tibble
homicide_test1 <- homicide_test %>% mutate(Black_majority = ifelse(max(Asian_Percentage_victim, Black_Percentage_victim, Native_Percentage_victim, Unspecified_Percentage_victim, White_Percentage_victim, na.rm = TRUE) == Black_Percentage_victim, 1, 0))

#checks for who many states was the model correct
sum(homicide_test1$Black_majority, na.rm = TRUE)

```
The base model is true for 21 states. Black is the most likely race of a homicide victim is 21 states even though only D.C has a black majority.




## Conclusion


The findings show that from 1990 to 2014, over 45% of the victims and 30% of the perpetrators of homicide crime in the U.S are Black even though the percentage of Black people in the U.S didn't exceed 15%, which demonstrates that Black communities are rife with crimes in contrast to White communities. Furthermore, in 21 states, the most likely race for the victim of Homicide is Black even though only D.C had a Black majority during the years of the study. In addition, over 20% of police shootings victims each year from 2000 to 2019 were Black, and the percentage was as high as 30% in 2012. In Rhode Island, the percentage of drivers stopped by the police who are Black are as high as double the proportion of Black people in the population of Rhode Island; furthermore, Black drivers were stopped for longer duration than non-Black drivers. For instance, 5% of the Black drivers were stopped by the police for more than 30 minutes compared to 3% of White drivers. These findings illustrate that Black people are targeted by the police at unjustified rates and underscores how Black communities are prone to the most dangerous crime, homicide. 

The analysis of the police stops in Rhode Island is limited in its span and can’t be generalized to all the United States. Also, one of my aims was to visualize the trend in racial bias in homicide, police shootings, and police stops across the time frame of each dataset. However, the years 1980 to 1989 in the homicide data set were omitted after joining the homicide and the Census Bureau data sets because the census data set didn’t have data for the U.S population from 1980 to 1990.  A next step might be working with more comprehensive data, such as a U.S census data set that includes all the years since 1980 and a data set of police stops across all the United States.